### 2020-12-05 doc/paper_ms_method_HSCY1/sparse_PointMass_results.tex

In this subsection, we show the results for the setup, which substitutes the
default NFW dictionary with the point mass dictionary, to compare with the
default setup.  The regularization parameter for the preliminary lasso is set
to $3.5$ and $5.0$.  Inspired by \citet{structureAdaLasso-Pramanik2020}, which
propose to incorporate external group information into different adaptive lasso
penalization weights by setting the penalization weights for projection
coefficients in the same group to the average of the adaptive weights in this
group, we smooth the preliminary lasso estimation in each lens redshift bin
with a top-hat filter of comoving scale $0.25~h^{-1}$ Mpc. The smoothed
preliminary lasso estimation is denoted as $\hat{x}^{\rm{ls}}_{\rm{sm}}$, and
the penalization weights are set to $1/\abs{\hat{x}^{\rm{ls}}_{\rm{sm}}}^\tau$.
The regularization parameter for the adaptive lasso is set to
$\lambda_{\rm{als}}=\lambda_{\rm{ls}}$.

As we have done on the NFW dictionary's reconstruction, we smooth the
reconstructed density maps with the Gaussian kernel (scale radius equals
$1.\arcmin 5$) in each lens redshift plane.

Figure \ref{fig_detFalsRatePM} shows the detection rate (left panels) and the
number of false peaks per square degree (right panel) for each setup of halo
mass and redshift.
This figure includes results of different penalization parameters:
$\lambda=3.5$ (upper panel) and $\lambda=5.0$ (lower panel). The right panels
of Figure \ref{fig_NFWvsPM} show the $3$-D density maps reconstructed with
different penalization parameters for a halo with $M_{200}=10^{15.02}
~h^{-1}M_{\odot}$ at redshift $0.164$.
Comparing with the results of the NFW dictionary, we find that the number of
false peaks in the mass map reconstructed with the point mass dictionary is
much larger than the NFW dictionary.

As demonstrated in the right panels of Figure \ref{fig_NFWvsPM}, the mass
reconstructions with the point mass dictionary tend to assign masses to
several different redshift bins in the neighboring region of the halo center.
While the NFW dictionary produces consistent mass reconstructions.
Since the NFW atoms' profiles are more consistent with the mass profiles of the
input halos than the profile of a point mass, there are less false structures
reconstructed with the NFW dictionary.