### 2020-12-07 doc/paper_ms_method_HSCY1/ms.tex
correct many typos

\documentclass[twocolumn]{aastex63}
\emergencystretch=1.em

\usepackage{amsmath,amsthm,amsfonts,amssymb,bm}
\usepackage{physics,commath}

\usepackage{listings}
\hypersetup{breaklinks}
\usepackage{color}
\usepackage{graphicx}
\usepackage{multirow}

\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{algorithm,algcompatible}
\usepackage[-]{callouts}

\makeatletter
\newcommand*\bigcdot{\mathpalette\bigcdot@{.5}}
\newcommand*\bigcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\newcommand{\argmax}{\mathop{\rm arg~max}\limits}
\newcommand{\argmin}{\mathop{\rm arg~min}\limits}

\DeclareMathOperator{\arccosh}{arccosh}
\newcommand \redColor{\color{red}}

\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi

\algnewcommand\INPUT{\item[\textbf{Input:}]}
\algnewcommand\OUTPUT{\item[\textbf{Output:}]}

\newcommand{\includezraphics}[1]{
    \begin{annotate}{
        \includegraphics[width=0.95\textwidth]{#1}
    }{1.0}
    \arrow { -3., -2}{ -3., 2}
    \note {-3.,2.5}{z}
    \end{annotate}
}

\makeatother

\begin{document}
\title{$3$-D Weak Lensing Mass Map Reconstruction:
Reducing the line of sight smearing}
% \author{Xiangchong Li}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \author{Masamune Oguri}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo
% 113-0033, Japan}
% \author{Wentao Luo}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from
photometric weak-lensing shear measurements. The $3$-D density contrast maps
are modeled as a summation of the NFW basis atoms, which have $2$-D multi-scale
NFW surface density profiles on the transverse plane and $1$-D Dirac delta
functions in the line of sight direction. With the prior assumption that the
density fields have a sparse spatial distribution, the density fields are
reconstructed using an oracle algorithm: adaptive lasso. The method is tested
with realistic simulations using the HSC-like shape estimation error and
photo-$z$ uncertainty.  Our findings are summarized as follows: 1) The lasso
solution suffers from a smear of structure in the line of sight direction even
in the absence of shape noise. In contrast, the adaptive lasso algorithm
efficiently removes the line of sight smear of structure.  2) The algorithm is
able to detect halo with minimal mass limits of $10^{14.0} M_{\odot}/h$,
$10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median
($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively, with a
false detection of 0.022/deg$^2$.  3) The estimated redshift of the halos
detected from the reconstructed mass maps are lower than the true redshift by
about $0.03$ for halos at low redshifts ($z\leq 0.4$). The relative redshift
bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

\section{Introduction}

\begin{figure*}[!t] \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis
        atoms in Fourier space, and the lower row shows the basis atoms in Real
        space.  The leftmost column is the point mass atom, and the other
        columns are the multi-scale NFW atoms.  The smoothing kernel is
        Gaussian with a standard dev iation of $1.5$ pixels.
    } \label{fig_atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.45\textwidth]{nfwlet-atom-1D.pdf}
    \caption{The $1$-D slices for smoothed pixelized basis atoms at $x=0$. The
        corresponding $2$-D profiles are shown in Figure \ref{fig_atoms2D}.
    }
 \label{fig_atoms1D}
\end{figure}

Weak lensing refers to the phenomenon that light from distant galaxies is
distorted by the intervening inhomogeneous density distribution along the line
of sight due to gravity's influence.
As a result of the light distortion, the shapes of the background galaxies are
coherently distorted. The lensing effect imprints the information of
foreground mass density distribution to the background galaxy images and offers
a direct probe into the mass density distribution in our universe
\citep[see][for recent reviews]{revKilbinger15,revRachel17}.

The expected shear measurements ($\gamma$) on distant galaxies are related to
the foreground density contrast field ($\delta$) through a linear
transformation: \begin{equation} \label{eq-intro-delta2shear} \gamma=\mathbf{T}
\delta, \end{equation} where $\mathbf{T}$ is used to denote the linear
transformation operator, which includes not only the physical lensing effect but
also systematic effects from observations (e.g., pixelization and smoothing of
the shear field in the transverse plane, photo-$z$ uncertainty).

Several large-scale surveys target to study the weak-lensing effect at a high
precision level (e.g., HSC \citep{HSC1-data}, KIDS \citep{KIDS13}, DES
\citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, NGRST
\citep{WFIRST15}).

The primary goal of most weak-lensing surveys is to constrain the cosmology
model through the $2$-point correlations. The studies include galaxy$-$galaxy
lensing, which cross-correlating the shear field ($\gamma$) with the positions
of foreground galaxies
\citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1}, and cosmic
shear which auto-correlates the shear measurements
\citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
Since the shear is directly related to the foreground matter distribution as
shown in eq. (\ref{eq-intro-delta2shear}), Galaxy$-$galaxy lensing probes into
the correlation between the matter field and galaxy field. On the other hand,
cosmic shear probes into the auto-correlation of matter field.

The reconstruction of density map from shear measurements also receive
considerable interest as it reaches the nonlinear scales. The $2$-D density map
reconstruction which recovers an integration of projected mass along the line of
sight has been well studied within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However,
the reconstruction of $3$-D mass map is still a challenging task.

To fully reconstruct the $3$-D mass density distribution ($\delta$) from the
photometric shear observations ($\gamma$), the density contrast field is
modeled as a summation of basis atoms in a model dictionary:
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection
coefficient field to the density contrast, and $x$ denotes the projection
coefficients. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density
field in Fourier space, which is equivalent to modeling the mass field with
sinusoidal functions.  On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} model the mass field with Starlets
\citep{Starlet-Starck2015}.

The projection coefficients are estimated by optimizing a regularized loss
function. The estimator is generally defined as
\begin{equation}
\hat{x}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma-
\mathbf{T\Phi} x)}_2^2+ \lambda C(x) \right\},
\end{equation}
where $\norm{\Sigma^{-\frac{1}{2}}(\gamma - \mathbf{T\Phi} x)}_2^2$ is the
$l^2$ chi-square term\footnote{weighted by the inverse of the diagonal
covariance matrix of the error on the shear measurements ($\Sigma$).} measuring
the difference between the prediction and the data, while $C(x)$ is the
regularization term measuring the deviation of the coefficient estimation ($x$)
from the prior assumption. The $l^p$ norm is defined as
\begin{equation}
\norm{x}_p=(\sum_{i} |x|^p_i)^{\frac{1}{p}}.
\end{equation}
Such penalized estimation prefers the parameters that are able to describe the
observations and align with the prior assumption. The regularization
parameter $\lambda$ adjusts the relative weight between the data and
prior assumption in the optimization process.

\citet{LSS-massMap-Wiener-Simon2009} propose to use the Wiener filter, which is
also known as $l^2$ ridge regulation ($C=\norm{x}^2_2$), to find a regularized
solution in Fourier space. \citet{HSC1-massMaps} apply the method of
\citet{LSS-massMap-Wiener-Simon2009} to the first-year data of the Hyper
Suprime-Cam Survey \citep{HSC1-data}.  However, the density maps reconstructed
by this method suffer from smearing along the line of sight direction with a
standard deviation of $\sigma_z=0.2 \sim 0.3$.

\citet{LSS-massMap-Glimpse3D-Leonard2014} propose the Glimpse algorithm, which
uses a derivative version of $l^1$ lasso regulation ($C=\norm{x}^1_1$) to find
a sparse solution in the Starlet dictionary space \citep{Starlet-Starck2015}.
\citet{LSS-massMap-Glimpse3D-Leonard2014} apply a greedy coordinate descent
algorithm, which selects the steepest coordinate in each iteration, to find the
minimum of a non-convex loss function penalized with the firm thresholding
function. The Glimpse algorithm reduces the smearing along the line of sight
since the coordinate descent algorithm forces the structure to grow only on the
most related lens redshift plane.  However, the stability of the non-convex
optimization and greedy coordinate descent algorithm has not been fully
justified. Moreover, the Starlet dictionary are not designed to model the
profile of clumpy mass in the universe, and the Starlet dictionary does not
account for the angular scale difference at different lens redshifts.

$N$-body simulations have shown that the dark matter is distributed in halos
connected by filaments, and the density profile of a single halo follows the
NFW function \citep{halo-NFW1997ApJ}.  We construct a model dictionary with the
multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles
of the NFW functions \citep{haloModel-TJ2003-3pt} on the transverse plane.
Following \citet{LSS-massMap-Glimpse3D-Leonard2014}, we neglect the depth of
halos since the resolution scale of the reconstruction in the line of sight
direction is much larger than the halo scale. Therefore, we set the NFW atoms'
profile in the line of sight direction as the Dirac delta function.  Moreover,
we assume that the halos are sparsely distributed in the universe.  With the
sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is
used to reconstruct the density field.  We find that, in contrast to the lasso
estimator that smears the structure along the line of sight, the adaptive lasso
can efficiently reduce the smearing effect.

Compared with \citet{LSS-massMap-Glimpse3D-Leonard2014}, our dictionary is
built up to describe the clumpy mass in the universe with a clear physical
motivation. Furthermore, the adaptive lasso algorithm is strictly convex and
can be directly optimized with the FISTA algorithm \citep{FISTA-Beck2009}
without relying on any greedy coordinate descent approaches. The stability of
this convex optimization has been justified.

This paper is organized as follows.
In Section \ref{sec_Method}, we propose the new method for $3$-D density map
reconstruction.
In Section \ref{sec_Test}, we test the novel algorithm on HSC-like simulations.
In Section \ref{sec_Sum}, we summarize and discuss the future development of
the method.

\section{Methodology}
\label{sec_Method}

\begin{figure}
 \centering
 \includegraphics[width=0.45\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to
     their Machine Learning and photo-Z (MLZ) best photo-$z$ estimation. The
     blue histogram is the normalized number distribution of the best photo-$z$
     estimation. The vertical dashed lines are the boundaries of the redshift
     bins.  The galaxies are equal-number binned.
        } \label{fig_bestpz}
\end{figure}

We first review the lensing effect in Section \ref{subsec_method_delta2shear}.
Then, we introduce the dictionary used to model the foreground density maps in
Section \ref{subsec_method_dictionary}.

Subsequently, in Section \ref{subsec_method_Systematics}, we discuss several
systematic effects from observations which include photo-$z$ uncertainty
(Section \ref{subsec_method_photoz}), smoothing (Section
\ref{subsec_method_smoothing}), masking(Section \ref{subsec_method_msknoise}),
and pixelization (Section \ref{subsec_method_pixel}).

Finally, we solve the mass reconstruction problem in Section
\ref{subsec_method_reconstruction} using the adaptive lasso algorithm
\citep{AdaLASSO-Zou2006} optimized with the FISTA algorithm
\citep{FISTA-Beck2009}.

\subsection{Lensing}
\label{subsec_method_delta2shear}

\begin{figure}
 \centering
 \includegraphics[width=0.45\textwidth]{mlz-poz.pdf}
 \caption{The average PDF of MLZ photo-$z$ error for $10$ source redshift bins.
        }\label{fig_pdfpz}
\end{figure}

The lensing convergence map at the comoving distance $\chi_s$ caused by the
foreground inhomogeneous density distribution at the comoving distance $\chi_l$
($\chi_l< \chi_s$) along the line of sight is
\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s}
\frac{\delta(\vec{\theta},\chi_l)}{a(\chi_l)},
\end{equation}
where $\delta=\rho(\vec{\theta},\chi_l)/\bar{\rho}-1$ is the density contrast
at the position of lens, $H_0$ is the Hubble parameter, $\Omega_M$ is the
matter density parameter, $c$ is the speed of light, and $a(\chi_l)$ is the
scale parameter at the lens position.

After substituting comoving distance ($\chi$) with redshift ($z$), we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l K(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
where $K(z_l,z_s)$ is the lensing kernel:
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l),
\end{cases}
\end{equation}
where $E(z)$ is the Hubble parameter as a function of redshift, in units of $H_0$.

\begin{figure*}[!t]
 \centering
 \includegraphics[width=1.\textwidth]{lensing_kernel.pdf}
 \caption{The left panel shows the lensing kernels for five different lens
         redshifts. The dashed lines are the kernels for spectroscopic redshift,
         which assumes that the source galaxies' redshifts are precisely
         estimated. The solid lines are for photometric redshifts, which
         accounts for the influence of photometric redshift uncertainty. The
         other two panels show the correlation between lensing kernels of
         different lens redshifts. The middle panel is for spectroscopic
         redshift, and the right panel is for photometric redshifts. The
         lensing kernels are normalized so that the diagonal elements of the
         correlation matrices equal one.
        }\label{fig_corlensKer}
\end{figure*}

As shown in \citet{massMap-KS1993}, the shear field is related to the kappa
field at the same source redshift plane via
\begin{equation}\label{eq-kappa2gamma}
\gamma_L(\vec{\theta},z_s) = \int  d^2 \theta' D(\vec{\theta}-\vec{\theta'}) \kappa(\vec{\theta'},z_s),
\end{equation}
where
\begin{equation}
D(\vec{\theta})=-\frac{1}{\pi}(\theta_1-i\theta_2)^{-2}.
\end{equation}
Here we denote the physical shear distortion as $\gamma_L$, and we note
that it is not the final shear measurement since the final shear
measurement is influenced by systematic errors from observations. The
systematic errors will be discussed in Section \ref{subsec_method_Systematics}.

Combining equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}),
the expectation of lensing shear signal is
\begin{equation}\label{eq-delta2gammat}
\gamma_L(\vec{\theta},z_s) = \int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}

To simplify the expression, we define the lensing transform operator as
\begin{equation}
\mathbf{Q}=\int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta'  \vec{D}(\vec{\theta}-\vec{\theta'}),
\end{equation}
and eq. (\ref{eq-delta2gammat}) is simplified to
\begin{equation} \label{eq-delta2gammat-simp}
\gamma_L=\mathbf{Q}\delta.
\end{equation}

\subsection{Dictionary}
\label{subsec_method_dictionary}

The density contrast field is modeled as a summation of basis atoms in the
dictionary:
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms
have `$N$' different scale frames, and the atoms in each scale frame are
shifted by $\vec{r'}$ to form models at different positions in the comoving
coordinates.  $x_s(\vec{r'})$ is the projection coefficient of the density
contrast field onto the basis atoms at the comoving coordinate: $\vec{r'}$.

We propose to use the multi-scale NFW atoms, denoted as
$\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary.  On the
transverse plane, the NFW atoms follow surface density profiles of the NFW
halos \citep{haloModel-TJ2003-3pt} with comoving scale radii $r_s$ and
truncation radius $c r_s$, where $c$ is the concentration of the NFW
halos.  As the scales of halos are much smaller than the reachable redshift
resolution, we neglect the depth of halo on the line of sight direction and set
the NFW atoms' profiles in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}. The
multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_s(\vec{r}_{\theta},z) =&\frac{f}{2 \pi r_s^2 }
F(|\vec{r}_{\theta}|/r_s) \delta_D(z),\\
&  (s=1...N)
\end{split}
\end{equation}
where $\vec{r}_\theta$ is the comoving position in the transverse plane,
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh
\left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} +
\frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in
different scale frames. Then, we write the basis atoms into the angular
separation coordinates:
\begin{equation}
\begin{split}
\phi_s(\vec{\theta},z) =&\frac{f \chi(z)^2}{2 \pi r_s^2 }
F(|\vec{\theta}|\chi(z)/r_s) \delta_D(z),\\
& (s=1...N)
\end{split}
\end{equation}
where $\chi(z)$ is the comoving distance at the redshift $z$.

To simplify the notation, we compress the projection coefficients into a column
vector:
$x=\begin{pmatrix}
x_{0}\\
x_{1}\\
...\\
x_{N}
\end{pmatrix}$,
and compress the dictionary transform operator to a row vector:
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
We substitute eq. (\ref{eq-x2delta}) into eq. (\ref{eq-delta2gammat}) and get
\begin{equation}\label{eq-x2gammat}
\gamma_L=\mathbf{Q}\mathbf{\Phi} x.
\end{equation}

In this paper, a dictionary constructed with point mass atoms is used to
compare with the dictionary of multi-scale atoms. The point mass atoms is a
$3$-D Dirac function defined as follows
\begin{equation}
\phi_{\rm{PM}}(\vec{\theta},z) = \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}

The $2$-D profiles of the point mass atom and the multi-scale NFW atoms on the
transverse plane are shown in Figure \ref{fig_atoms2D}. The $1$-D slices of the
profiles are demonstrated in Figure \ref{fig_atoms1D}. Note that these profiles
are smoothed with a Gaussian kernel and pixelized into evenly spaced grids. The
smoothing operation is discussed in Section \ref{subsec_method_smoothing}, and
the pixelization operation is discussed in Section \ref{subsec_method_pixel}.

\subsection{Systematics}
\label{subsec_method_Systematics}

The observed shear measurement is deviated from the physical shear prediction
due to the systematic errors from the observation. The influence of systematics is
carefully studied and incorporated into the forward modeling in this subsection.

\subsubsection{Photo-z Uncertainty}
\label{subsec_method_photoz}

\begin{figure*}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-lasso.pdf}
    \centering
    \small lasso
\end{minipage}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-alasso.pdf}
    \centering
    \small adaptive lasso
\end{minipage}
\caption{The density map reconstructions with the lasso (left) and the adaptive
    lasso (right) algorithms. The vertical direction is the line of sight
    direction, and the lower boundaries and upper boundaries of the boxes are
    $z=0.01$ and $z=0.85$, respectively. Noises on galaxy shape measurements
    are neglected in the simulation.  The input halo mass is $M_{200}=10^{15}
    ~h^{-1}M_{\odot}$, and its redshift is $z=0.35$. The lasso reconstruction
    smears the density field along the line of sight direction.
    } \label{fig_lassoVsadaLasso}
\end{figure*}

The photometric redshifts of source galaxies in the current large-scale survey
are estimated with a limited number of broad photometric bands (e.g., $9$ bands
for KIDS$+$VIKING survey \citep{KIDS_VIKING-Hildebrant2020}, $5$ bands for DES
survey and HSC survey). As a result, the estimated redshifts of galaxies suffer
from much larger uncertainty than the redshifts estimated with spectroscopic
observations. Such photo-$z$ uncertainty smears the lensing kernels
statistically since a galaxy with a best-fit photo-$z$ estimation of $z_s$ has
possibilities of being actually located at different redshifts ($z$). The
probability function is denoted as $P(z|z_s)$,
and the expected shear distortion on the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator:
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
we express the influence of photo-$z$ uncertainty on the shear signal as
\begin{equation}
\gamma_L \rightarrow \mathbf{P} \gamma_L.
\end{equation}

Figure \ref{fig_bestpz} shows the histogram of the best Machine Learning and
photo-Z \citep[MLZ]{MLZ-TPZ2013} photometric redshift estimation from
\cite{HSC1-photoz} for galaxies in the tract 9347 of the HSC S16A data release
\citep{HSC1-data}. These Galaxies are divided into ten source galaxy bins
according to the photo-$z$ best-fit estimation, and the boundaries of the bins are
shown as vertical dashed lines in Figure \ref{fig_bestpz}. Figure \ref{fig_pdfpz}
shows the average probability density function (PDF) for galaxies in each
redshift bin.

The left panel of Figure \ref{fig_corlensKer} shows the lensing kernels for
lenses at five different redshifts as functions of source galaxy redshifts.
The dashed lines are the lensing kernels for spectroscopic redshifts with
neglectable redshift uncertainties. The solid lines are the lensing kernels for
photometric redshifts with redshift uncertainties shown in Figure
\ref{fig_pdfpz}. As shown by the dashed lines in the left panel of Figure
\ref{fig_corlensKer}, the lensing kernels converge to zero for source redshifts
lower than the lens redshift if the uncertainty on the source galaxy
redshift estimation are neglectable. However, as demonstrated by the solid
lines in the same panel, for the source redshifts with large photo-$z$
uncertainties, the lensing kernels do not converge to zero at redshifts lower
than the lens redshift. This is because the galaxies with photo-$z$ estimations
lower than the lens redshifts may be actually located at higher redshifts due
to the photo-$z$ uncertainties.

The middle and right panels show the correlations between lensing kernels for
lenses at different redshifts. The middle panel is for spect-$z$, and the right
panel is for photo-$z$. As demonstrated in the middle panel, the lensing
kernels are highly correlated even though the redshift estimation is precise.
Comparing the correlation matrices shown in the middle panel and the right
panel of Figure \ref{fig_corlensKer}, we conclude that the photo-$z$
uncertainty slightly increases the correlations between lensing kernels at
different lens plane.

\subsubsection{Smoothing}
\label{subsec_method_smoothing}

The observed galaxies have a random yet unequally-spaced spatial distribution
in the transverse plane. To boost the computational speed, we smooth the shear
measurements from galaxies and pixelize the smoothed shear field onto a regular
grid. After the pixelization, the fast Fourier transform (FFT) can be directly
conducted on the transverse plane in each source redshift bin.  Another benefit
of smoothing is that it reduces the bias from the aliasing effect in the
pixelization process as the smoothing kernel reduces the amplitude of the shear
signal at the high frequency.

The smoothing is conducted by convolving the shear measurements with a
smoothing kernel:
\begin{equation}
\gamma_{\rm{sm}} (\vec{\theta})  = \frac{\sum_i
W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) \gamma_i}{\sum_i
W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $\gamma_i$, $z_i$ and
$\theta_i$ are the shear, photometric reshift, and transverse position of the
$i$-th galaxy in the catalog.

$W(\vec{\theta},z)$ can be decomposed into a transverse component
$W_T(\vec{\theta})$ and a line of sight component
$W_\times(z)$ as
\begin{equation}
W(\vec{\theta},z)=W_T(\vec{\theta}) W_\times (z).
\end{equation}
In this paper, we use an isotropic $2$-D Gaussian kernel and a $1$-D top-hat
kernel to smooth the measurements in the transverse plane and the line of sight
direction. These components of the smoothing kernel are
\begin{equation}
\begin{split}
W_T(\vec{\theta}) &=\frac{1}{2\pi\beta^2}\exp(-\frac{|\vec{\theta}|}{2\beta^2}),\\
W_\times (z) &=
\begin{cases}
1/\Delta z& (|z|<\Delta z/2),\\
0& else.
\end{cases}
\end{split}
\end{equation}
In this paper, we set $\beta=1.\arcmin5$.

By definition, the smoothing kernel is normalized as
\begin{equation}
\int d^3r W(\vec{r})=1.
\end{equation}
With the approximation that the density of galaxy number - $n(\vec{r})$ -
varies slowly at the smoothing scale, the smoothed galaxy number density,
which is defined as
\begin{equation}\label{eq-numDesSmooth}
n_{\rm{sm}}(\vec{r})=\sum_i W(\vec{\theta}-\vec{\theta_i},z-z_i),
\end{equation}
equals the galaxy number density: $n_{\rm{sm}}(\vec{r})=n(\vec{r})$.
However, the galaxy number density experience a steep drop on the boundary
of the survey, therefore the smoothed galaxy number density does not equal the
galaxy number density close to the boundary of the survey.

The smoothing operator is defined as
\begin{equation}
\mathbf{W} = \int d^3 r' W(\vec{r}-\vec{r'}),
\end{equation}
and the smoothing procedure influence the shear signal by
\begin{equation}
\gamma_L \rightarrow \mathbf{W} \gamma_L.
\end{equation}

As we will discuss in Section \ref{subsec_method_pixel}, the smoothed shear
field is pixelized into equally spaced grids. We note that another widely used
scheme is to average the shear measurements in each pixel.  Such a scenario is
equivalent to resampling the shear field smoothed with a $3$-D top-hat kernel
with the same scale as the pixels.

\subsubsection{Masking}
\label{subsec_method_msknoise}

In real observations, shear measurements are available in a finite region of
the sky, and the boundary of the region is always irregular. Moreover, many
isolated sub-regions near the bright stars are masked out since the light from
bright stars tends to influence the shear measurements on neighboring galaxies.

We define the masking window function according to the smoothed galaxy number
density (defined in eq. (\ref{eq-numDesSmooth})):
\begin{equation}
 M(\vec{r})=
\begin{cases}
0 & n_{\rm{sm}}>1,\\
1 & \rm{else}.
\end{cases}
\end{equation}
The mask changes the shear measurements by
\begin{equation}\label{eq-delta2gamma-final}
\gamma_L(\vec{\theta},z) \rightarrow M(\vec{\theta},z) \gamma_L(\vec{\theta},z),
\end{equation}
We define the masking operator as
\begin{equation}
\mathbf{M}= \int d^3 r' M(\vec{r'}) \delta_D(\vec{r}-\vec{r'}),
\end{equation}
where $\delta_D(\vec{r})$ is $3$-D Dirac delta function. The shear is
influenced by the masking by
$\gamma_L \rightarrow \mathbf{M} \gamma_L$.

The final observed shear field, taking into account all of the systematics as
mentioned above from observations, is
\begin{equation}\label{eq-x2gamma-final}
\gamma =\mathbf{M} \mathbf{W} \mathbf{P} \mathbf{Q} \mathbf{\Phi} x.
\end{equation}
For simplicity, we denote $\mathbf{A}=\mathbf{M} \mathbf{W} \mathbf{P}
\mathbf{Q} \mathbf{\Phi} $ and
eq. (\ref{eq-x2gamma-final}) is written as
\begin{equation}\label{eq-x2gamma-simple}
\gamma=\mathbf{A} x.
\end{equation}

\subsubsection{Pixelization}
\label{subsec_method_pixel}

We pixelize the smoothed shear field into an $N_\theta \times N_\theta \times
N_s$ grid, where $N_\theta$ is the number of pixels for the two orthogonal axes
of the transverse plane and $N_s$ is the number of pixels for the line of sight
axis. $\gamma_{\alpha}$ denotes the smoothed shear measurements recorded on the
pixel with index $\alpha$, where $\alpha=1...N_\theta \times N_\theta \times
N_s$. The grids on the transverse planes are equally spaced with a pixel size
of $1\arcmin$. While the grids in the line of sight direction follow
equal number binning as shown in Figure \ref{fig_bestpz}.

Similarly, we pixelize each scale frame of the projection coefficient field $x$
into an $N_\theta \times N_\theta \times N_l$ grid. The pixelization on the
transverse plane for each scale frame is the same as that of the
smoothed shear field on the transverse plane. At the same time, the projection
coefficient field is pixelized into equal spaced grids in the line of sight
direction ranging from redshift $0.01$ to redshift $0.85$. Here, we use $N_l$
to denote the number of the lens planes and $x_{\beta}$ to denote the
projection parameter indexed as $\beta$, where $\beta=1...N_\theta \times
N_\theta \times N_l \times N$. The corresponding pixelized elements of the
forward transform matrix $\mathbf{A}$ is denoted as $A_{\alpha\beta}$.

We term the column vectors of the forward transform matrix $\mathbf{A}$ as the
effective basis atoms. The $l^2$ norm of the $i$-th column vectors weighted by
the inverse of the noise covariance matrix's diagonal elements is defined as
\begin{equation}
\mathcal{N}_{i}=\sum_\alpha A_{i\alpha}A_{i\alpha}/\Sigma_{\alpha\alpha}.
\end{equation}

We note that the effective basis atoms have different weighted $l^2$ norm.
Before solving the density map reconstruction problem, we normalize the column
vectors of the transform matrix through a rescaling:
\begin{equation}
\begin{split}
A'_{\alpha\beta}&=A_{\alpha\beta}/\mathcal{N}_{\alpha}^{\frac{1}{2}},\\
x'_{\beta}&=x_{\beta}\mathcal{N}_{\beta}^{\frac{1}{2}},\\
\end{split}
\end{equation}
to make the projection coefficients have the same weighted $l^2$ norm.
Such normalization boosts the speed of the gradient descent iterations
in the next subsection.

\subsection{Density map reconstruction}
\label{subsec_method_reconstruction}

\subsubsection{Adaptive lasso}

The lasso algorithm uses $l^1$ norm of the projection coefficient field to
regularize the modeling, and the estimator is defined as
\begin{equation}
\hat{x'}^{\rm{lasso}}=\argmin_{x} \left\{
\frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{A'} x')}_2^2+
\lambda \norm{x'}^1_1\right\},
\end{equation}
where $\norm{\bigcdot}_1$ and $\norm{\bigcdot}_2$ refer to the $l^1$ norm and
$l^2$ norm, respectively.  $\lambda$ denotes the penalization parameter
for the lasso estimation.

The lasso algorithm selects the parameters relevant to the measurements and
simultaneously estimates the value of the selected parameters.  However, it has
been shown by \citet{AdaLASSO-Zou2006} that when the column vectors of the
transforming matrix $\mathbf{A'}$ are highly correlated, the lasso cannot
select the relevant parameters from the parameter space consistently.
Moreover, the estimated parameters are biased due to the shrinkage in the lasso
regression. We note that, for the density map reconstruction problem, the
column vectors are highly correlated even in the absence of photometric
redshift uncertainties because, as shown in Figure \ref{fig_corlensKer}, the
lensing kernels for lenses at different redshifts are highly correlated.
Therefore, the lasso algorithm cannot select the consistent mass in the line of
sight direction, and the reconstructed mass suffers from smearing in the line
of sight direction even in the absence of noise on shear measurements and
photo-$z$ uncertainty .

Figure \ref{fig_lassoVsadaLasso} shows the reconstructions of a single halo's
mass map with halo mass equals $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift
$0.35$. The shear measurement error and photo-$z$ uncertainty are not included
in the simulation.  The left panel of Figure \ref{fig_lassoVsadaLasso} is the
reconstruction with the lasso algorithm, which shows a significant smear along
the line of sight.

\citet{AdaLASSO-Zou2006} proposes the adaptive lasso algorithm, which uses
adaptive weights to penalize different projection coefficients in the $l^1$
penalty. The adaptive lasso algorithm is a two-steps process. In the first
step, the lasso is used to estimate the parameters, and the preliminary
estimation of the lasso is denoted as $\hat{x'}^{\rm{lasso}}$. In the second
step, the preliminary lasso estimation is used to weight the penalization. The
weight on penalty is defined as
\begin{equation}
\hat{w}= \frac{1}{\abs{\hat{x'}^{\rm{lasso}}}^\tau},
\end{equation}
where we set the hyper-parameter $\tau$ to $2$.  The adaptive lasso estimator is
expressed as
\begin{equation}\label{eq-lossFun}
\hat{x'}=\argmin_{x'} \left\{ \frac{1}{2} \norm{\Sigma^{-\frac{1}{2}}(\gamma-\mathbf{A'}x')}^2_2 +
\hat{w}\lambda_{\rm{ada}} \norm{x'}^1_1\right\}.
\end{equation}
Here $\lambda_{\rm{ada}}$ is the penalization parameter for the adaptive lasso,
which does not need to be the same as the penalization parameter for the
preliminary lasso estimation ($\lambda$).

We rewrite the loss function with the Einstein notation:
\begin{equation}
\begin{split}
L(x')&=\frac{1}{2}(\Sigma^{-1})_{\alpha\beta}(\gamma^{*}_{\alpha}-A'^{*}_{\alpha i}x'_{i})
(\gamma_{\beta}-A'_{\beta j}x'_{j})\\
&+ \lambda_{\rm{ada}} \hat{w_\beta} \abs{x'_\beta}.
\end{split}
\end{equation}
To simplify the notification in future, we define the quadruple term in the
loss function as $G(x')$:
\begin{equation}
G(x')=\frac{1}{2}\Sigma^{-1}_{\alpha\beta}(\gamma^{*}_{\alpha}-A'^{*}_{\alpha
i}x'_{i}) (\gamma_{\beta}-A'_{\beta j}x'_{j}).
\end{equation}

\subsubsection{FISTA}

\citet{FISTA-Beck2009} propose the Fast Iterative Soft Thresholding Algorithm
(FISTA) to solve the lasso problem.  Since the lasso's loss function and the
adaptive lasso's loss function only differ in their penalization terms, the
FISTA is also applicable to solve the adaptive lasso problem. In this paper, we
apply the FISTA to solve both the preliminary lasso estimation and the final
adaptive lasso estimation.

Here we start from the lasso preliminary estimation. The coefficients are
initialized as $x_i^{(1)}=0$. According to the FISTA algorithm, we iteratively
update the projection coefficient field ($x$). Taking the $n$'th iteration as
an example, a temporary update is first calculated as
\begin{equation}
x'^{(n+1)}_{i}=\mathrm{ST}_{\lambda} \left(x'^{(n)}_{i} -\mu \partial_i G(x'^{(n)})\right),
\end{equation}
where $\mathrm{ST}$ is the soft thresholding function defined as
\begin{equation}
\mathrm{ST}_{\lambda} \left(x'\right) = \mathrm{sign} (x') \max \left(\abs{x'}-\lambda,0\right).
\end{equation}
The soft thresholding is a part of the lasso algorithm. It selects the
modes with an amplitude greater than $\lambda$, and shrink the selected
estimations by
$\lambda$.
$\mu$ is the step size of the gradient descent iteration.  $\partial_i
G(x'^{(n)})$ refers to the $i$'th element of the gradient vector of $G$ at
point $x'^{(n)}$:
\begin{equation}
\partial_i G(x'^{(n)})=\Sigma^{-1}_{\alpha\beta}\Re\left(A'^{*}_{\alpha i}(\gamma_{\beta}-A'_{\beta j}x'_{j})\right),
\end{equation}
where $\Re\left( \bullet \right)$ is the function returns the real part of the
input function. The FISTA algorithm requires an additional update amounting to a
weighted average between
$x'^{(n+1)}$ and $x'^{(n)}$:
\begin{equation}
\begin{split}
t^{(n+1)}&=\frac{1+\sqrt{1+4(t^{(n)})^2}}{2},\\
x'^{(n+1)} &\leftarrow x'^{(x+1)}+ \frac{t^{(n)}-1}{t^{(n+1)}}(x'^{(n+1)}-x'^{(n)}),
\end{split}
\end{equation}
where the relative weight is initialized as $t^{(1)}=1$.

The FISTA algorithm converges as long as the gradient descent step size $\mu$
satisfies
\begin{equation}
 0< \mu < \frac{1}{ \norm{\mathbf{A^{\dagger}} \mathbf{\Sigma}^{-1} \mathbf{A} }},
\end{equation}
where $\norm{\mathbf{A^{\dagger}} \mathbf{\Sigma}^{-1} \mathbf{A} }$ refers to
the spectrum norm of the matrix $\mathbf{A^{\dagger}} \mathbf{\Sigma}^{-1}
\mathbf{A}$. The spectral norm is estimated by simulating a large number of
random vectors with $l^2$ norms equal one with different realizations. The
matrix operator $\mathbf{A^{\dagger}} \mathbf{\Sigma}^{-1} \mathbf{A}$ is
subsequently applied to each random vector and get a corresponding transformed
vector. The spectral norm of the matrix $\mathbf{A^{\dagger}}
\mathbf{\Sigma}^{-1} \mathbf{A}$ is determined as the maximum $l^2$ norm of the
transformed vectors.

As summarized in Algorithm \ref{alg-1}, we first initial the projection
coefficients as zero and use the FISTA algorithm to find the global minimum of
the lasso loss function. Such a global minimum is the preliminary estimation.
We then use the preliminary lasso estimation to weight the coefficients and
construct the adaptive lasso loss function. Finally, we set the preliminary
lasso estimation as the warm start of the adaptive lasso estimation and use the
FISTA algorithm again to find the adaptive lasso loss function's global minimum,
which is the final solution.

\begin{algorithm}[H]
\renewcommand{\thealgorithm}{}
\label{alg-1}
\caption{Our Algorithm}
\begin{algorithmic}[1]
\INPUT $\gamma$: Pixelized complex $3$-D array of shear
\OUTPUT  $\delta$: $3$-D array of density contrast
\STATE Normalize column vectors of $\mathbf{A}$
\STATE Estimate step size $\mu$ and $\mathbf{\Sigma}$
\STATE \textbf{Initialization:}
\STATE $x'^{(1)} = 0$
\STATE $\hat{w}=1$
\STATE $t^{(1)}=1$,$i=1$, $j=1$
\WHILE{$j \leq 2$}
    \WHILE{$i \leq N_{\rm{iter}}$}
        \STATE \# soft thresholding
        \STATE $x'^{(n+1)}_{i}=\mathrm{ST}_{\hat{w}\lambda} \left(x'^{(n)}_{i} -\mu \partial_i G(x'^{(n)})\right)$
        \STATE \# FISTA algorithm
        \STATE $t^{(n+1)}=\frac{1+\sqrt{1+4(t^{(n)})^2}}{2}$
        \STATE $x'^{(n+1)} \leftarrow x'^{(x+1)}+ \frac{t^{(n)}-1}{t^{(n+1)}}(x'^{(n+1)}-x'^{(n)})$
        \STATE $i=i+1$
    \ENDWHILE
\STATE \textbf{Reinitialization:}
\STATE $\hat{w}=\abs{\hat{x'}^{\rm{lasso}}}^{-2}$, $\lambda \leftarrow
\lambda_{\rm{ada}}$
\STATE $\hat{x'}^{(1)} = x'^{(N_{\rm{iter}})}$
\STATE $t^{(1)}=1$, $i=1$
\STATE $j=j+1$
\ENDWHILE
\STATE $\delta=\mathbf{\Phi}\mathcal{N}^{-\frac{1}{2}}x'^{(N_{\rm{iter}})}$
\end{algorithmic}
\end{algorithm}

\section{Tests}
\label{sec_Test}

This Section simulates weak-lensing shear fields induced by a group of NFW
halos with various halo masses and redshifts. The shear fields are used to
distort the HSC mock shape catalogs with different realizations of the HSC-like
shape measurement error and photo-$z$ uncertainty (Section \ref{subsec_Sims}).

Then, we test our algorithm using the simulations with different setups of the
regularization parameter. We also compare the results of our algorithm, which
uses the NFW dictionary (Section \ref{subsec_test-nfw}), with the point mass
dictionary (Section \ref{subsec_test-pm}).

\subsection{Simulations}
\label{subsec_Sims}
\input{massmap_sim.tex}

\subsection{NFW atoms}
\label{subsec_test-nfw}
\input{sparse_NFW_results.tex}

\subsection{Point mass atoms}
\label{subsec_test-pm}
\input{sparse_PointMass_results.tex}

\section{Summary}
\label{sec_Sum}

We develop a novel method to reconstruct $3$-D density contrast maps from
weak-lensing shear measurements and photometric redshift estimations.  Our
method models $3$-D density contrast maps as summations of NFW atoms with
different comoving radii.  With the prior assumption that the clumpy masses
sparsely distribute in the $3$-D space, the density map is reconstructed
using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}. The method is
tested with realistic simulations using HSC-like shape estimation error and
photo-$z$ uncertainty.

The findings of this paper are summarized as follows:

(i) The lasso algorithm's solution suffers from a smear of structure in the
line of sight direction even in the absence of shape noise, and the adaptive
lasso algorithm efficiently removes the line of sight smear of structure.

(ii) The algorithm is able to detect halo with minimal mass limits of
$10^{14.0} M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for
the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$)
redshifts, respectively, with a false detection of 0.022/deg$^2$.

(iii)The estimated redshifts of the halos detected from the reconstructed mass
maps are lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$).  The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$.

We will apply the method to the shear measurements of the HSC survey
\citep{HSC1-catalog,FPFSHSC1-Li2020} to perform galaxy cluster detection in our
future work.

\section*{Acknowledgements}
XL thank Yin Li and Jiaxin Han for useful discussions.

XL was supported by Global Science Graduate Course (GSGC) program of
University of Tokyo and JSPS KAKENHI (JP19J22222).
\bibliographystyle{aasjournal}
\bibliography{citation}

\end{document}