### 2021-01-13 doc/paper_ms_method_HSCY1/ms.tex

\documentclass[twocolumn]{aastex63}
\emergencystretch=1.em

\usepackage{amsmath,amsthm,amsfonts,amssymb,bm}
\usepackage{physics,commath}

\usepackage{listings}
\hypersetup{breaklinks}
\usepackage{color}
\usepackage{graphicx}
\usepackage{multirow}

\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{algorithm,algcompatible}
\usepackage[-]{callouts}

\makeatletter
\newcommand*\bigcdot{\mathpalette\bigcdot@{.5}}
\newcommand*\bigcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\newcommand{\argmax}{\mathop{\rm arg~max}\limits}
\newcommand{\argmin}{\mathop{\rm arg~min}\limits}

\DeclareMathOperator{\arccosh}{arccosh}
\newcommand \redColor{\color{red}}

\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi

\algnewcommand\INPUT{\item[\textbf{Input:}]}
\algnewcommand\OUTPUT{\item[\textbf{Output:}]}

\newcommand{\includezraphics}[1]{
    \begin{annotate}{
        \includegraphics[width=0.95\textwidth]{#1}
    }{1.0}
    \arrow { -3., -2}{ -3., 2}
    \note {-3.,2.5}{z}
    \end{annotate}
}

\makeatother

\begin{document}
\title{A new method cluster Detection from three-dimensional weak lensing mass map}
% \author{Xiangchong Li}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \author{Masamune Oguri}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo
% 113-0033, Japan}
% \author{Wentao Luo}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct high-resoluion three-dimensional mass
maps from photometric weak-lensing shear measurements. The three-dimensional
mass map is modeled as a summation of the NFW basis atoms, which have
two-dimensional multi-scale NFW surface density profiles on the transverse
plane and one-dimensional Dirac delta functions in the line of sight direction.
The adaptive lasso algorithm is applied to find a sparse reconstruction of the
mass maps.
We study the performance of three-dimensional detection of clusters of galaxies
from the reconstructed mass map with simulations that apply shear distortions
from isolated halo to HSC-like galaxy shapes with realistic photometric
redshift uncertainties .
We find that the adaptive lasso algorithm efficiently removes the line of sight
smearing, in contrast to the lasso reconstruction that suffers from  the line
of sight smearing even in the absence of shape noise.
The algorithm is able to detect halo with minimal mass limits of $10^{14.0}
h^{-1}M_{\odot}$, $10^{14.7} h^{-1}M_{\odot}$, $10^{15.0} h^{-1}M_{\odot}$ for the low
($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts,
respectively, with an average false detection of 0.022/deg$^2$.
The estimated redshift of the halos detected from the reconstructed mass maps
are slightly lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$). The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$. The standard deviation of the redshift estimation is
$0.092$.
\end{abstract}

\section{Introduction}

\begin{figure*}[!t] \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis
        atoms in Fourier space, and the lower row shows the basis atoms in Real
        space.  The leftmost column is the point mass atom, and the other
        columns are the multi-scale NFW atoms.  The smoothing kernel is
        Gaussian with a standard deviation of $1.5$ pixels.
    } \label{fig_atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.45\textwidth]{nfwlet-atom-1D.pdf}
    \caption{The $1$-D slices for smoothed pixelized basis atoms at $x=0$. The
        corresponding $2$-D profiles are shown in Figure \ref{fig_atoms2D}.
    }
 \label{fig_atoms1D}
\end{figure}

Weak lensing refers to the phenomenon that light from distant galaxies is
weakly yet coherently distorted by the intervening inhomogeneous density
distribution along the line of sight due to the influence of gravity.  As a
result, the shapes of the background galaxies are distorted such that the
information of the foreground mass distribution is imprinted to the background
galaxy images. Weak lensing offers a direct probe into the mass distribution,
including both visible matter and invisible dark matter, in our universe
\citep[see][for recent reviews]{revKilbinger15,revRachel17}.
Several large-scale surveys focus on studying the weak-lensing effect at a high
precision level e.g., HSC \citep{HSC1-data}, KIDS \citep{KIDS13}, DES
\citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, NGRST
\citep{WFIRST15}.

% The primary goal of most weak-lensing surveys is to constrain the cosmology
% model through the $2$-point correlations. The studies include galaxy$-$galaxy
% lensing, which cross-correlating the shear field ($\gamma$) with the positions
% of foreground galaxies
% \citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1}, and cosmic
% shear which auto-correlates the shear measurements
% \citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
% Since the shear is directly related to the foreground matter distribution as
% shown in eq. (\ref{eq-intro-delta2shear}), Galaxy$-$galaxy lensing probes into
% the correlation between the matter field and galaxy field. On the other hand,
% cosmic shear probes into the auto-correlation of matter field.

The reconstruction of density map from shear measurements has received
considerable interest since
we can constrain the cosmology model using the peak counts, which are
identified from the weak lensing mass map
\citep{WL-massMap-peakcounts-Jain2000,WL-massMap-peakcountsAna-Fan2010,WL-massMap-peakcountsFM-Lin2016}.
Moreover, we can directly detect massive clusters of galaxies through
identifying high signal-to-noise (SNR) peaks on the mass map without any
reference to the mass-to-light ratio
\citep{WL-massMap-clusDet-Schneider1996,WL-massMap-clusDet-Hamana2004}.

The two-dimensional ($2$-D) density map reconstruction which recovers an
integration of projected mass along the line of sight has been well studied
within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}.
Cluster detection from $2$-D mass maps has been widely studied and applied to
these large-scale weak-lensing surveys
\citep{WL-massMap-clusDet-CFHT-Shan2012,WL-massMap-clusDet-HSC-Miyazaki2018,WL-massMap-clusDet-HSC-Hamana2020};
however, cross matching with some (e.g., optically selected) cluster catalog is
needed to assign the redshift and extract its physical parameter (such as mass)
for each cluster selected from $2$-D mass maps.

On the other hand, we can directly reconstruct three-dimensional ($3$-D) mass
maps taking advantage of photometric redshift (photo-$z$) information
\citep{LSS-massMap-Wiener-Simon2009,WL-massMap-VanderPlas2011}; however, these
methods either do not have enough spatial resolution to identify clusters, or
they suffer from smearing along the line of sight direction, which need to be
overcome for practical searches of clusters in 3D mass maps.  Alternatively,
\citet{WL-clusDet-Hennawi2005} propose to perform a maximum-likelihood
detection of clusters, by convolving tomographic shear measurements with $3$-D
filters that match the tangential shears induced by multi-scale NFW halos,
without fully reconstruct the $3$-D mass maps.

In this paper, we develop a novel method for high-resolution $3$-D mass map
reconstruction that are free from line of sight smearing.  The $3$-D
mass map is modeled as a summation of the NFW \citep{halo-NFW1997ApJ} basis
atoms, which have $2$-D multi-scale NFW surface density profiles on the
transverse plane and one-dimensional Dirac delta functions in the line of sight
direction. The adaptive lasso algorithm \citep{AdaLASSO-Zou2006} is applied to
find a sparse reconstruction of the mass maps.

We study the performance of cluster detection from the reconstructed mass maps
using simulations applying shear distortions from isolated halos to HSC-like
galaxy shapes with realistic photo-$z$ uncertainties.

This paper is organized as follows.
In Section \ref{sec_Method}, we propose the new method for $3$-D density map
reconstruction.
In Section \ref{sec_Test}, we study the cluster detection from the
reconstructed mass map using isolated halo simulations with the HSC
observational condition.
In Section \ref{sec_Sum}, we summarize and discuss the future development of
the method.

\section{$3$-D Mass Map Reconstruction}
\label{sec_Method}
\input{massmap_method.tex}

\section{Cluster Detection}
\label{sec_Test}

This Section simulates weak-lensing shear fields induced by a group of NFW
halos with various halo masses and redshifts. The shear fields are used to
distort the HSC mock shape catalogs with different realizations of the HSC-like
shape measurement error and photo-$z$ uncertainty (Section \ref{subsec_Sims}).

We then test our algorithm using the simulations with different setups of the
regularization parameter. We also compare the results of our algorithm, which
uses the NFW dictionary (Section \ref{subsec_test-nfw}), with the point mass
dictionary (Section \ref{subsec_test-pm}).

\subsection{Simulations}
\label{subsec_Sims}
\input{massmap_sim.tex}

\subsection{NFW atoms}
\label{subsec_test-nfw}
\input{sparse_NFW_results.tex}

\subsection{Point mass atoms}
\label{subsec_test-pm}
\input{sparse_PointMass_results.tex}

\section{Summary}
\label{sec_Sum}

We have developed a novel method to reconstruct high-resolution $3$-D density
contrast maps from weak-lensing shear measurements and photometric redshift
estimations.  Our method models $3$-D density contrast maps as summations of
NFW atoms with different comoving radii.  With the prior assumption that the
clumpy masses sparsely distribute in the $3$-D space, the density map is
reconstructed using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}.
We find that the lasso algorithm uffers from smearing of structure
in the line of sight direction even in the absence of galaxy shape noise, and
the adaptive lasso algorithm efficiently removes the line of sight smearing of
structure.

We study the performance of cluster detection from the reconstructed $3$-D mass
maps using simulations that apply shear distortions from isolated NFW halos to
HSC-like shapes and photo-$z$ uncertainties.  We conclude that the algorithm is
able to detect halo with minimal mass limits of $10^{14.0} h^{-1}M_{\odot}$,
$10^{14.7} h^{-1}M_{\odot}$, $10^{15.0} h^{-1}M_{\odot}$ for the low ($z<0.3$),
median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively,
with an average false detection of 0.022 deg$^{-2}$. The estimated redshifts of
the halos detected from the reconstructed mass maps are slightly lower than the
true redshift by about $0.03$ for halos at low redshifts ($z\leq 0.4$).  The
relative redshift bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$. The
standard deviation of the redshift estimation is $0.092$.

We will apply the $3$-D mass map reconstruction to the shear measurements of
the HSC survey \citep[e.g.,][]{HSC1-catalog,FPFSHSC1-Li2020} to perform galaxy
cluster detection in our future work.

\section*{Acknowledgements}
We thank Yin Li and Jiaxin Han for useful discussions.

XL was supported by Global Science Graduate Course (GSGC) program of University
of Tokyo and JSPS KAKENHI (JP19J22222).

This work was supported in part by Japan Science and Technology Agency (JST)
CREST JPMHCR1414, and by JST AIP Acceleration Research Grant No. JP20317829,
Japan. This work was supported in part by the World Premier International
Research Center Initiative (WPI Initiative), MEXT, Japan, and JSPS KAKENHI
Grant Nos. JP18K03693, JP20H00181, JP20H05856.

\bibliographystyle{aasjournal}
\bibliography{citation}

\end{document}